# Chapter 2 {#chap2}

Here is another chapter...

## Creating figures 

```{r foo, fig.cap='testing figure captions.', fig.align = 'center'}
plot(cars)
```

Can cross-ref your figures, see Figure \@ref(fig:foo).^[You can also use footnotes in your book.]

Block quotes are a thing:

> "I thoroughly disapprove of duels. If a man should challenge me,
  I would take him kindly and forgivingly by the hand and lead him
  to a quiet place and kill him."
>
> --- Mark Twain

```{r irisPlot, fig.cap='Scatterplot with legend using iris data.', fig.align = 'center'}
plot(iris$Sepal.Length, iris$Petal.Length, col = iris$Species, pch = 16, cex = 2)
legend(4.5, 7, 
       legend = c("setosa","versicolor","virginica"),
       col = c(1:3),
       lty = "solid")
```

## Simulate data and plot some figures

```{r simData}
## Simulate some data

## 3 Factor Variables
FacVar1 = as.factor(rep(c("level1", "level2"), 25))
FacVar2 = as.factor(rep(c("levelA", "levelB", "levelC"), 17)[-51])
FacVar3 = as.factor(rep(c("levelI", "levelII", "levelIII", "levelIV"), 13)[-c(51:52)])

## 4 Numeric Vars
set.seed(123)
NumVar1 = round(rnorm(n = 50, mean = 1000, sd = 50), digits = 2)  ## Normal distribution
set.seed(123)
NumVar2 = round(runif(n = 50, min = 500, max = 1500), digits = 2)  ## Uniform distribution
set.seed(123)
NumVar3 = round(rexp(n = 50, rate = 0.001))  ## Exponential distribution
NumVar4 = 2001:2050

simData = data.frame(FacVar1, FacVar2, FacVar3, NumVar1, NumVar2, NumVar3, NumVar4)
```

### One Variable: Numeric Variable

```{r oneVar, fig.cap='Plotting one numeric variable.', fig.align='center'}
plot(simData$NumVar1, type = "o")  ## Index plot
```

### Two Variables: Two Numeric Variables

```{r twoVar, fig.cap='Plotting two numeric variables.', fig.align='center'}
plot(simData$NumVar1, type = "o", ylim = c(0, max(simData$NumVar1, simData$NumVar2)))  ## index plot with one variable
lines(simData$NumVar2, type = "o", lty = 2, col = "red")  ## add another variable
```


