# Chapter 2 {#chap2}

Here is another chapter...

## Creating figures 

```{r foo, fig.cap='testing figure captions.', fig.align = 'center'}
plot(cars)
```

Can cross-ref your figures, see Figure \@ref(fig:foo).^[You can also use footnotes in your book.]

Block quotes are a thing:

> "I thoroughly disapprove of duels. If a man should challenge me,
  I would take him kindly and forgivingly by the hand and lead him
  to a quiet place and kill him."
>
> --- Mark Twain

```{r irisPlot, fig.cap='Scatterplot with legend using iris data.', fig.align = 'center'}
plot(iris$Sepal.Length, iris$Petal.Length, col = iris$Species, pch = 16, cex = 2)
legend(4.5, 7, 
       legend = c("setosa","versicolor","virginica"),
       col = c(1:3),
       lty = "solid")
```

## Simulate data and plot some figures

```{r simData}
## Simulate some data

## 3 Factor Variables
FacVar1 = as.factor(rep(c("level1", "level2"), 25))
FacVar2 = as.factor(rep(c("levelA", "levelB", "levelC"), 17)[-51])
FacVar3 = as.factor(rep(c("levelI", "levelII", "levelIII", "levelIV"), 13)[-c(51:52)])

## 4 Numeric Vars
set.seed(123)
NumVar1 = round(rnorm(n = 50, mean = 1000, sd = 50), digits = 2)  ## Normal distribution
set.seed(123)
NumVar2 = round(runif(n = 50, min = 500, max = 1500), digits = 2)  ## Uniform distribution
set.seed(123)
NumVar3 = round(rexp(n = 50, rate = 0.001))  ## Exponential distribution
NumVar4 = 2001:2050

simData = data.frame(FacVar1, FacVar2, FacVar3, NumVar1, NumVar2, NumVar3, NumVar4)
```

### One Variable: Numeric Variable

```{r oneVar, fig.cap='Plotting one numeric variable.', fig.align='center'}
plot(simData$NumVar1, type = "o")  ## Index plot
```

### Two Variables: Two Numeric Variables

```{r twoVar, fig.cap='Plotting two numeric variables.', fig.align='center'}
plot(simData$NumVar1, type = "o", ylim = c(0, max(simData$NumVar1, simData$NumVar2)))  ## index plot with one variable
lines(simData$NumVar2, type = "o", lty = 2, col = "red")  ## add another variable
```

## More plotting

The below examples come from here: http://shinyapps.org/apps/RGraphCompendium/index.php

### Correlations

```{r}
# Presidential data up to and including 2008; data from Stulp et al. 2013
# rm(list=ls())
# height of president divided by height of most successful opponent: 
height.ratio <- c(0.924324324, 1.081871345, 1, 0.971098266, 1.029761905,
   0.935135135, 0.994252874, 0.908163265, 1.045714286, 1.18404908,
   1.115606936, 0.971910112, 0.97752809, 0.978609626, 1,
   0.933333333, 1.071428571, 0.944444444, 0.944444444, 1.017142857,
   1.011111111, 1.011235955, 1.011235955, 1.089285714, 0.988888889,
   1.011111111, 1.032967033, 1.044444444, 1, 1.086705202,
   1.011560694, 1.005617978, 1.005617978, 1.005494505, 1.072222222,
   1.011111111, 0.983783784, 0.967213115, 1.04519774, 1.027777778,
   1.086705202, 1, 1.005347594, 0.983783784, 0.943005181, 1.057142857)

# proportion popular vote for president vs most successful opponent
# NB can be lower than .5 because popolar vote does not decide election
pop.vote <- c(0.427780852, 0.56148981, 0.597141922, 0.581254292, 0.530344067,
  0.507425996, 0.526679292, 0.536690951, 0.577825976, 0.573225387,
  0.550410082, 0.559380032, 0.484823958, 0.500466176, 0.502934212,
  0.49569636, 0.516904414, 0.522050547, 0.531494442, 0.60014892, 
  0.545079801, 0.604274986, 0.51635906, 0.63850958, 0.652184407, 
  0.587920412, 0.5914898, 0.624614752, 0.550040193, 0.537771958, 
  0.523673642, 0.554517134, 0.577511576, 0.500856251, 0.613444534, 
  0.504063153, 0.617883695, 0.51049949, 0.553073235, 0.59166415, 
  0.538982024, 0.53455133, 0.547304058, 0.497350649, 0.512424242, 
  0.536914796)
           
# cor.test(height.ratio,pop.vote)
library(plotrix) # package plotrix is needed for function "ablineclip""
# if the following line and the line containing "dev.off()" are executed, the plot will be saved as a png file in the current working directory
# png("Presidental.png", width = 18, height = 18, units = "cm", res = 800, pointsize = 10) 
op <- par(cex.main = 1.5, mar = c(5, 6, 4, 5) + 0.1, mgp = c(3.5, 1, 0), cex.lab = 1.5 , font.lab = 2, cex.axis = 1.3, bty = "n", las = 1)
plot(height.ratio, pop.vote, col = "black", pch = 21, bg = "grey", cex = 2,
     xlim = c(.90,1.20), ylim = c(.40,.70), ylab = "", xlab = "", axes = FALSE)
axis(1)
axis(2) 
reg1 <- lm(pop.vote ~ height.ratio)
ablineclip(reg1, lwd = 2,x1 = .9, x2 = 1.2) 
par(las = 0)
mtext("Presidential Height Ratio", side = 1, line = 2.5, cex = 1.5)
mtext("Relative Support for President", side = 2, line = 3.7, cex = 1.5)
text(1.15, .65, "r = .39", cex = 1.5)
# dev.off()
# For comparison, consider the default plot:
par(op) # reset to default "par" settings
plot(height.ratio, pop.vote) #yuk!
```

### Including a Density Estimator

```{r}
# rm(list = ls())
# Data: Proportion of choices from the good decks as reported in 39 studies
good.choices <- c(.43, .47, .47, .48, .50, .52, .53, .53, .54, .54, .54, .54, .55, .55, .55, .56, .56, .57, .57, .57, .57, .58, .58, .58, .59, .59, .60, .62, .63, .63, .64, .64, .66, .66, .67, .67, .68, .70, .70)
yhigh <- 8
par(cex.main = 1.5, mar = c(5, 6, 4, 5) + 0.1, mgp = c(3.5, 1, 0), cex.lab = 1.5 , font.lab = 2, cex.axis = 1.3, bty = "n", las = 1)
h <- hist(good.choices, freq = FALSE, main = "", xlab = "", ylab = " ", ylim = c(0, yhigh), xlim = c(.30, .80), axes = FALSE, col = "grey")
axis(1, seq(.30, .80, by = .1))
axis(2, labels = FALSE, lwd.ticks = 0)
rug(jitter(good.choices))
mtext("Prop. Choices from Good Decks", side = 1, line = 2.5, cex = 1.5, font = 2)
mtext("Density of Studies", side = 2, line = 1, cex = 1.5, font = 2, las = 0)
lines(density(good.choices), lwd = 2)
```

### Network graph

```{r, message=F}
library("psych")
library("qgraph")
```

```{r network, message=F, fig.align='center', fig.width=15}
# Load BFI data:
data(bfi)
bfi <- bfi[, 1:25]

# Groups and names object (not needed really, but make the plots easier to
# interpret):
Names <- scan("http://sachaepskamp.com/files/BFIitems.txt", what = "character", sep = "\n")

# Create groups object:
Groups <- rep(c("A", "C", "E", "N", "O"), each = 5)

# Compute correlations:
cor_bfi <- cor_auto(bfi)

# Plot correlation network:
graph_cor <- qgraph(cor_bfi, layout = "spring", nodeNames = Names, groups = Groups, legend.cex = 0.6, 
    DoNotPlot = TRUE)

# Plot partial correlation network:
graph_pcor <- qgraph(cor_bfi, graph = "concentration", layout = "spring", nodeNames = Names, 
    groups = Groups, legend.cex = 0.6, DoNotPlot = TRUE)

# Plot glasso network:
graph_glas <- qgraph(cor_bfi, graph = "glasso", sampleSize = nrow(bfi), layout = "spring", 
    nodeNames = Names, legend.cex = 0.6, groups = Groups, legend.cex = 0.7, GLratio = 2)
```
