# Part II: Experimental Structure {-}

# An example experiment write up {#writeup}

This chapter will detail the structure that an experimental write up should take for this course.

What follows is a template for the basic structure laid out in the WILS Lab Book pdf manual.

## Title
**Using R Notebooks for electronic lab book documentation in Bioinformatics** (and generally any other field you want!)

## Aims
Explore the use of the R Notebook environment for creating a reproducible and dynamic reporting system for use in the field of Bioinformatics.

## Introduction
It has been well established that the tools contained within the RStudio software have made creating and diseminating electronic reports much easier. By using RMarkdown users are able to generate notes/written content in the same ernvironment as working with their data and producing statistics and results. 

## Materials
For this demonstration we need to download and install several pieces of software. 

First would be `R`, which can be downloaded here: https://www.r-project.org/  

Next we'll want to grab `RStudio`, here is the link: https://www.rstudio.com/products/rstudio/download/preview/

If you want to read more about RMarkdown: http://rmarkdown.rstudio.com/

More information on R Notebooks: http://rmarkdown.rstudio.com/r_notebooks.html

## Methods
We will be exploring a range of different methods and functions from both R and RStudio, and see how they are implemented in R Notebooks and how that can be leveraged to help us create an electronic record, or lab book.

## Results
Here are some of the many features that are available when using R Notebooks. 

### Display code {#codeDisplay}
We are able to write and evaluate code using *chunks*, a concept implemented in RMarkdown - allowing us to both write text and code within the same document.

To genetate a new chunk press *ctrl+alt+i*, within the chunk you can write and evaluate R code:
```{r}
1+1
```

### Evaluate statistics/models and display in text

```{r}
lm(cars)
```

```{r}
summary(lm(cars))
```

### Can even include plots  

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Ctrl+Shift+Enter*. 

```{r mtcars, fig.cap='Simple plotting using mtcars data.', tidy=TRUE, fig.align = 'center'}
plot(cars, pch = 16, col = 'cadetblue')
```

A really nifty feature is that plotted figures get asigned a figure number which is able to be referenced in text (see Figure \@ref(fig:mtcars)).

We can create the same plot as before, but this time fit the regression model that we previously generated. We'll plot the best fit line using the *abline* function.

```{r regplot, fig.cap='Plot with regression model fitted.', fig.align = 'center'}
plot(cars, pch = 16, col = 'cadetblue')
abline(lm(cars$dist ~ cars$speed))
```

And again we can easily cross-ref the above plot, Figure \@ref(fig:regplot).

### Tables are also easy
We can use the *kable* function from *knitr* (note that tables can also be captioned):
```{r kable}
require(knitr)
n <- 100
x <- rnorm(n)
y <- 2*x + rnorm(n)
out <- lm(y ~ x)
kable(summary(out)$coef, digits=2, caption = 'An example table produced by knitr kable().')
```

There is a new package called *printr* which is a companion app to *knitr*, aiming to make tables look nicer.

The package is currently not on CRAN, so you will need to install it as below:
```{r, eval=FALSE}
# install printr
install.packages(
  'printr',
  type = 'source',
  repos = c('http://yihui.name/xran', 'http://cran.rstudio.com')
)
# load the package
library('printr')
```

The *printr* package aims to make the standard output from `R` a little nicer to look at (and easier to read), have a look at the following:

```{r}
library('printr')
head(mtcars)
```

### Output to many popular formats
RStudio uses the power of the `knitr` package and `pandoc` to allow the exporting of your document to several flexible and popular formats.

#### HTML
By selecting *knit to HTML* an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Ctrl+Shift+K* to preview the HTML file).

#### Word (.docx)
By selecting *knit to Word* your notebook will be exported to *.docx* format, which will be able to be read by Microsoft Word and other office packages (i.e. Libre Office).

#### pdf
By selecting *knit to Word* your notebook will be exported to *.pdf*, the universal format.

## Discussion
For any heavily based computational or 'informatics program we should be moving towards embracing technology not only in the tools we use for research, but also those for methods of recording and diseminating our work. As researchers in this field are almost exclusively working on their computers it makes sense to have an electronic record rather thantraditional hand written lab books.

This set up is able to run on any OS platform (Linux, MacOS, Windows, etc).

## Conclusion
The framework provided by R, RStudio and RMarkdown come together in the form of R Notebooks and make for an excellent form of electronic recording for daily activities/experiments, providing an ideal solution for Bioinformatics lab books.  

